---
title: "SABER melanoma data analysis"
output: html_document
---
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
```

## 2. Clustering 2. general cell type 
# clustering
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
selected_markers <- c('CD3','CD4','CD11c','CD20','CD15','MPO','SMA','CD31','Caveolin','HLA-DR', 'SOX9','MelanA','gp100', "S100","CD68","CD206")

pg_res <- Rphenograph(data = t(assay(sub[selected_markers,],"exprs")), k = 10)
sub$PG <- factor(pg_res[[2]]$membership)
```

```{r}
qual_col_pals <- brewer.pal.info[brewer.pal.info$category == 'qual',]   
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
pie(rep(1,length(col_vector)), col=col_vector, labels = names(col_vector))

an <- list()
an_col <- list()
ts_pal<-list()
# ts_pal$rygbw<-c("#F47C7C","#F7F48B","#A1DE93","#70A1D7","gray")
# ts_pal$dark<-c("darkred","goldenrod","darkgreen","navy","gray")
ts_pal$def<-c("firebrick2","gold","limegreen","navy","gray")
ts_pal$green<-c("darkgreen","yellowgreen","yellow2","gray")
ts_pal$red<-c("gold","darkorange","darkred","gray")
pie(rep(1,12),col=c(ts_pal$def,ts_pal$green,ts_pal$red))
# ts_pal$blue<-c("palegreen1","seagreen3","darkgreen","cyan3","darkcyan","powderblue","lightslateblue","purple2","mediumorchid1","darkred")

ts_pal$blue<-c(c(brewer.pal(9,name="Greens")[c(3,5,8)],brewer.pal(9,name="Blues")[c(3,8)],brewer.pal(9,name="Reds")[c(2,4,5,7,9)]))
pie(rep(1,10),col=c(ts_pal$blue))

```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(sub)), sub$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(sub[selected_markers,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(sub, "TSNE") <- matrix(NA, ncol(sub), ncol(xy)) #create empty template
reducedDim(sub, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```


```{r, fig.width=8}
DR <- "TSNE"
p <- .get_df(sub[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =6) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'RdYlBu')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```
```{r, fig.width=6}
DR <- "TSNE"
var <- "PG"
p <- .get_df(sub[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```
#preprocess for heatmap, median expression per PG cluster
#plot heatmap for check
```{r warning = FALSE, fig.width=8, fig.height=5}
hm_dat <- .get_matrix(sub, group_by="PG")

groups =4
#distance = "correlation" #euclidean
distance = "euclidean"
method = "ward.D2"

ph <- pheatmap(hm_dat[,selected_markers],cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method=method)

ph_allm <- pheatmap(hm_dat,cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method=method)

# fn <- sprintf("%s_bigC_PG_heatmap_%s.png", prefix, suffix)
# fn_allm <- sprintf("%s_bigC_PG_heatmap_allmarkers_%s.png", prefix, suffix)
#  # ggsave(file.path(results_folder,  fn), ph, width = 6, height = 8)
#  # ggsave(file.path(results_folder,  fn_allm), ph_allm, width = 6, height = 8)
```

#plot heatmap
```{r, fig.width=8, fig.height=8}
an$pg <- data.frame(cluster=rownames(hm_dat))  %>% 
  mutate(annotation = ifelse(cluster%in% c(24,20,22,25,3,4,9,14,15,18,21), "Immune",
                               ifelse(cluster%in% c(13,23), "Vascular",
                               ifelse(cluster%in% c(7,8,17,12), "Fibroblasts",
                               ifelse(cluster%in% c(11,27,28), "Unassigned","Tumor")))))

an_col$pg <- list(Cell_type=c("Immune"=ts_pal$def[1],
                              "Vascular"=ts_pal$def[2],
                              "Fibroblasts"=ts_pal$def[3],
                              "Tumor"=ts_pal$def[4],
                              "Unassigned"=ts_pal$def[5]))

ph2 <- pheatmap(hm_dat[,selected_markers],cutree_rows = 1, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="complete", 
  border_color=NA, annotation_row = data.frame(Cell_type=an$pg$annotation), show_rownames = F, cellwidth = 20, cellheight = 10, 
  annotation_colors = an_col$pg, fontsize = 20, legend_breaks = seq(0.1,1,by=0.2))

fn2 <- sprintf("%s_heatmap_2_general_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width = 10, height = 10)
```

#assigning to PGa
```{r}
m <- match(sub$PG, an$pg$cluster)
sub$PGa <- an$pg$annotation[m]
table(sub$PGa, sub$sample_ROI_id)
```

```{r, fig.width=6}
DR <- "TSNE"
var <- "PGa"
p <- .get_df(sub[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```
## 3. Clustering 3. Immune cell type
## Re-clustering on PG-assigned "Immune" cells
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
tmp <- sub[,sub$PGa %in% c("Immune")]
selected_markers <- c('CD3','CD4','CD8a','CD11c','CD68','CD20','CD45RA','CD11b','CD15','MPO','HLA-DR', "CD206")

pg_res <- Rphenograph(data = t(assay(tmp[selected_markers,],"exprs")), k = 10)
tmp$PGi <- factor(pg_res[[2]]$membership)

sub$PGi <- 0
sub[,colnames(tmp)]$PGi <- tmp$PGi
unique(sub$PGi)
```

```{r}
unique(tmp$PGi)
```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(tmp)), tmp$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(tmp[selected_markers,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(tmp, "TSNE") <- matrix(NA, ncol(tmp), ncol(xy)) #create empty template
reducedDim(tmp, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```


```{r, fig.width=8}
DR <- "TSNE"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =6) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'RdYlBu')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```

```{r, fig.width=6}
DR <- "TSNE"
var <- "PGi"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=2)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```


#plot heatmap (for check)
```{r warning = FALSE, fig.width=8, fig.height=5}

hm_dat <- .get_matrix(tmp, group_by="PGi")
groups =4
#distance = "correlation" #euclidean
distance = "euclidean"

ph <- pheatmap(hm_dat[,selected_markers],cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward.D2")

ph_allm <- pheatmap(hm_dat,cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward.D2")

# fn <- sprintf("%s_immune_PG_heatmap_%s.png", prefix, suffix)
# fn_allm <- sprintf("%s_immune_PG_heatmap_allmarkers_%s.png", prefix, suffix)
#  ggsave(file.path(results_folder,  fn), ph, width = 6, height = 8)
#  ggsave(file.path(results_folder,  fn_allm), ph_allm, width = 6, height = 8)

```
```{r warning = FALSE, fig.width=8, fig.height=8}


an$pgi <- data.frame(cluster=rownames(hm_dat))  %>% 
  mutate(annotation = ifelse(cluster%in% c(1,6,3,20,11,12,5,29), "T-cells",
                               ifelse(cluster%in%c(32,30,31,28,25,26), "B-cells",
                               ifelse(cluster%in% c(10,15,13,7), "Unassigned",
                               ifelse(cluster%in% c(14), "Granulocytes", "Macrophages")))))

an_col$pgi <- list(Cell_type=c("Macrophages"=ts_pal$def[1],
                              "T-cells"=ts_pal$def[2],
                              "B-cells"=ts_pal$def[3],
                              "Granulocytes"=ts_pal$def[4],
                              "Unassigned"=ts_pal$def[5]))

ph2 <- pheatmap(hm_dat[,selected_markers],cutree_rows = 1, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="complete", border_color=NA, 
  annotation_row = data.frame(Cell_type=an$pgi$annotation), show_rownames = F, cellwidth = 20, cellheight = 10, 
  annotation_colors = an_col$pgi, fontsize = 20, legend_breaks = seq(0.1,1,by=0.2))
fn2 <- sprintf("%s_heatmap_3_immune_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width = 8, height = 8)
```
#assigning  to PGia
```{r}
m <- match(tmp$PGi, an$pgi$cluster)
tmp$PGia <- an$pgi$annotation[m]
sub$PGia <- sub$PGa
sub[,tmp$id]$PGia <- tmp$PGia
table(sub$PGia)
```

```{r, fig.width=5}
DR <- "TSNE"
var <- "PGia"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```
#plot frequency
```{r}
#per sample


cell_colors <-col_vector[1:length(unique(tmp$PGia))]
names(cell_colors) <- unique(tmp$PGia)

ns <- table(tmp$sample_ROI_id_fig, tmp$PGia)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(ROI=Var1, cell=Var2)

  p <- ggplot(soi_cf, aes(ROI, Freq, fill = cell)) +
    scale_fill_manual("cell type", values = cell_colors, drop=FALSE) +
    geom_bar(stat = "identity") + ggtitle('') + theme_light()
  p

  #save the plot
  # fn <- sprintf("%s_frequency_3_immune_ROI_%s.pdf", prefix,suffix)
  # ggsave(file.path(results_folder, fn), p, width = 12, height = 4)  
# per ROI
ns <- table(tmp$sample_ROI_dsc_fig, tmp$PGia)
#re-order
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(Sample=Var1, cell=Var2)
soi_cf$cell<-factor(soi_cf$cell, levels=c("Macrophages","T-cells","B-cells","Granulocytes","Unassigned"))

 
p <- ggplot(soi_cf, aes(Sample,Freq, fill = cell)) +
    scale_fill_manual("Cell type", values = ts_pal$def[c(1,2,3,4,5)], drop=FALSE) +geom_bar(stat = "identity") + 
  ggtitle('') + theme_light(base_size = 16)+
  theme(legend.title = element_text(face = "bold")) + labs(y ="Frequency")+
  scale_x_discrete(labels=c("Excluded", "Inflamed","Mixed\nexcluded","Mixed\ninflamed"))
  
p

  #save the plot
  fn <- sprintf("%s_frequency_3_immune_sample_%s.pdf", prefix,suffix)
  ggsave(file.path(results_folder, fn), p, width = 6, height = 4)  
```


## 4. Clustering 4. Macrophages
#PG clustering on "Macrophages" from 3.immune PG
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
tmp <- sub[,sub$PGia %in% "Macrophages"]
selected_markers <- c('HLA-DR','VISTA','CD11b','CD68','TIM3','CD11c','PDL1','CD86','PDL2','Axl','CD14','CD206' )
pg_res <- Rphenograph(data = t(assay(tmp[selected_markers,],"exprs")), k = 100)

tmp$PGm <- factor(pg_res[[2]]$membership, )

sub$PGm <- 0
sub[,colnames(tmp)]$PGm <- tmp$PGm
unique(sub$PGm)

```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(tmp)), tmp$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(tmp[selected_markers,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(tmp, "TSNE") <- matrix(NA, ncol(tmp), ncol(xy)) #create empty template
reducedDim(tmp, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```


```{r, fig.width=8}
DR <- "TSNE"
var <- "scaled"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =6) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```

```{r, fig.width=8}
DR <- "TSNE"
var <- "PGm"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```


#preprocess for heatmap, median expression per PG cluster
```{r}
hm_dat <- .get_matrix(tmp, group_by="PGm")
```

#plot heatmap (for check)
```{r warning = FALSE, fig.width=8, fig.height=5}
groups =3
#distance = "correlation" #euclidean
distance = "euclidean"

ph <- pheatmap(hm_dat[,selected_markers],cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward.D2")

ph_allm <- pheatmap(hm_dat,cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward.D2")

# fn <- sprintf("%s_macro_PG_heatmap_%s.png", prefix, suffix)
# fn_allm <- sprintf("%s_macro_PG_heatmap_allmarkers_%s.png", prefix, suffix)
#  ggsave(file.path(results_folder,  fn), ph, width = 6, height = 8)
#  ggsave(file.path(results_folder,  fn_allm), ph_allm, width = 6, height = 8)

```
#plot heatmap 
```{r warning = FALSE, fig.width=8, fig.height=8}


an$pgm <- data.frame(cluster=rownames(hm_dat))  %>% 
  mutate(annotation = ifelse(cluster%in% c(2), "PDL1/CD86_ISM",
                               ifelse(cluster%in%c(10), "CD68_ISM",
                               ifelse(cluster%in% c(3), "CD11c/HLA-DR_ISM", "Other_macrophages"))))

an_col$pgm <- list(Cell_type=c("PDL1/CD86_ISM"=ts_pal$green[3],
                              "CD68_ISM"=ts_pal$green[2],
                              "CD11c/HLA-DR_ISM"=ts_pal$green[1],
                              "Other_macrophages"=ts_pal$green[4]))

ph2 <- pheatmap(hm_dat[,selected_markers],cutree_rows = 1, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="complete", 
  border_color=NA, annotation_row = data.frame(Cell_type=an$pgm$annotation), show_rownames = F, cellwidth = 20, cellheight = 15, 
  annotation_colors = an_col$pgm, fontsize = 16, legend_breaks = seq(0,1,by=0.2))

fn2 <- sprintf("%s_heatmap_4_macrophages_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width =8 , height = 6)
```
#assigning  to PGma
```{r}
m <- match(tmp$PGm, an$pgm$cluster)
tmp$PGma <- an$pgm$annotation[m]
sub$PGma <- sub$PGa
sub[,tmp$id]$PGma <- tmp$PGma
table(sub$PGma)
```


#plot frequency
```{r}
#per ROI
ns <- table(tmp$sample_ROI_id_fig, tmp$PGma)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(ROI=Var1, cell=Var2)

  p <- ggplot(soi_cf, aes(ROI, Freq, fill = cell)) +
    scale_fill_manual("cell type", values = col_vector, drop=FALSE) +
    geom_bar(stat = "identity") + ggtitle('') + theme_light()
  p

  #save the plot
  # fn <- sprintf("%s_frequency_4_macrophage_ROI_%s.pdf", prefix,suffix)
  # ggsave(file.path(results_folder, fn), p, width = 12, height = 4)  

#per sample

ns <- table(tmp$sample_ROI_dsc_fig, tmp$PGma)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(Sample=Var1, cell=Var2)
#re-order
soi_cf$cell<-factor(soi_cf$cell, levels=c("Other_macrophages","PDL1/CD86_ISM","CD11c/HLA-DR_ISM","CD68_ISM"))

 
p <- ggplot(soi_cf, aes(Sample,Freq, fill = cell)) +
    scale_fill_manual("Cell type", values = ts_pal$green[c(4,3,2,1)], drop=FALSE) +geom_bar(stat = "identity") + 
  ggtitle('') + theme_light(base_size = 16)+
  theme(legend.title = element_text(face = "bold")) + labs(y ="Frequency")+
  scale_x_discrete(labels=c("Excluded", "Inflamed","Mixed\nexcluded","Mixed\ninflamed"))
  
p

  #save the plot
  fn <- sprintf("%s_frequency_4_macrophage_Sample_%s.pdf", prefix,suffix)
  ggsave(file.path(results_folder, fn), p, width = 7, height = 4)  

```


## 5. Clustering 5. T cells
#PG clustering on "T-cells" from 3.immune PG
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
tmp <- sub[,sub$PGia %in% "T-cells"]
selected_markers <-  c('CD4','CD8a','CTLA4','PD-1','TIM3','CD28','FoxP3','LAG3' )
pg_res <- Rphenograph(data = t(assay(tmp[selected_markers,],"exprs")), k = 80)
tmp$PGt <- factor(pg_res[[2]]$membership)
sub$PGt <- 0
sub[,colnames(tmp)]$PGt <- tmp$PGt
unique(sub$PGt)
```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(tmp)), tmp$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(tmp[selected_markers,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(tmp, "TSNE") <- matrix(NA, ncol(tmp), ncol(xy)) #create empty template
reducedDim(tmp, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```


```{r, fig.width=8}
DR <- "TSNE"
var <- "scaled"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =6) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'RdYlBu')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```

```{r, fig.width=8}
DR <- "TSNE"
var <- "PGt"
p <- .get_df(tmp[selected_markers,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```

#preprocess for heatmap, median expression per PG cluster




#plot heatmap 

```{r warning = FALSE, fig.width=8, fig.height=5}

hm_dat <- .get_matrix(tmp,group_by="PGt")
pheatmap(hm_dat)

groups =4
#distance = "correlation" #euclidean
distance = "euclidean"

ph <- pheatmap(hm_dat[,selected_markers],cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward")

# fn <- sprintf("%s_T-cells_PG_heatmap_%s.png", prefix, suffix)
# fn_allm <- sprintf("%s_T-cells_PG_heatmap_allmarkers_%s.png", prefix, suffix)
#  ggsave(file.path(results_folder,  fn), ph, width = 6, height = 8)
#  ggsave(file.path(results_folder,  fn_allm), ph_allm, width = 6, height = 8)

```
#plot heatmap 
```{r warning = FALSE, fig.width=8, fig.height=8}


an$pgt <- data.frame(cluster=rownames(hm_dat))  %>% 
  mutate(annotation = ifelse(cluster%in% c(3,6), "Exhausted_CD8",
                               ifelse(cluster%in%c(2), "CTLA4_CD4",
                               ifelse(cluster%in% c(4), "PD1_CD4", "Other_T-cells"))))

an_col$pgt <- list(Cell_type=c("Exhausted_CD8"=ts_pal$red[1],
                              "CTLA4_CD4"=ts_pal$red[2],
                              "PD1_CD4"=ts_pal$red[3],
                              "Other_T-cells"=ts_pal$red[4]))

ph2 <- pheatmap(hm_dat[,selected_markers],cutree_rows = 1, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="complete", 
  border_color=NA, annotation_row = data.frame(Cell_type=an$pgt$annotation), show_rownames = F, cellwidth = 20, cellheight = 15, 
  annotation_colors = an_col$pgt, fontsize = 16, legend_breaks = seq(0,1,by=0.2))

fn2 <- sprintf("%s_heatmap_5_T-cells_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width =8 , height = 6)
```
#assigning  to PGta
```{r}
m <- match(tmp$PGt, an$pgt$cluster)
tmp$PGta <- an$pgt$annotation[m]
sub$PGta <- sub$PGa
sub[,tmp$id]$PGta <- tmp$PGta
table(sub$PGta)
```
#plot frequency
```{r}
ns <- table(tmp$sample_ROI_id_fig, tmp$PGt)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(ROI=Var1, cell=Var2)

  p <- ggplot(soi_cf, aes(ROI, Freq, fill = cell)) +
    scale_fill_manual("cell type", values = col_vector, drop=FALSE) +
    geom_bar(stat = "identity") + ggtitle('') + theme_light()
  p

  #save the plot
  # fn <- sprintf("%s_frequency_5_T-cells_ROI_%s.pdf", prefix,suffix)
  # ggsave(file.path(results_folder, fn), p, width = 12, height = 4)  

ns <- table(tmp$sample_ROI_dsc_fig, tmp$PGta)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(Sample=Var1, cell=Var2)
#re-order
soi_cf$cell<-factor(soi_cf$cell, levels=c("Other_T-cells","Exhausted_CD8","CTLA4_CD4","PD1_CD4"))

 
p <- ggplot(soi_cf, aes(Sample,Freq, fill = cell)) +
    scale_fill_manual("Cell type", values = ts_pal$red[c(4,1,2,3)], drop=FALSE) +geom_bar(stat = "identity") + 
  ggtitle('') + theme_light(base_size = 16)+
  theme(legend.title = element_text(face = "bold")) + labs(y ="Frequency")+
  scale_x_discrete(labels=c("Excluded", "Inflamed","Mixed\nexcluded","Mixed\ninflamed"))
  
p

  #save the plot
  fn <- sprintf("%s_frequency_5_T-cells_Sample_%s.pdf", prefix,suffix)
  ggsave(file.path(results_folder, fn), p, width = 6.5, height = 4)  

  
```

```{r}
ns <- table(tmp$sample_ROI_dsc_fig, tmp$PGt)
df <- as.data.frame(ns)%>%rename(Sample=Var1, cell_type=Var2, Count=Freq)

  
  p <- ggplot(df, aes(Sample, Count, fill=cell_type)) +
    scale_fill_manual("cell_type", values = col_vector, drop=FALSE) +
    geom_bar(position = "dodge",stat = "identity") + ggtitle('') + theme_light()
  p
```



##6. Clustering 6. Tumor cells
# clustering on PG-defiened Tumor cells

```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
tmp <- sub[,sub$PGa %in% "Tumor"]
selected_markers <- c('HLA-ABC','gp100','S100','Glut1','PDL1','SOX10','TRP1','Ki-67','SOX9','MelanA' )
pg_res <- Rphenograph(data = t(assay(tmp[selected_markers,],"exprs")), k = 80)
tmp$PGtu <- factor(pg_res[[2]]$membership)
sub$PGtu <- 0
sub[,colnames(tmp)]$PGtu <- tmp$PGtu
unique(sub$PGtu)

```

```{r}
hm_dat <- .get_matrix(tmp,group_by="PGtu")
```

#plot heatmap
```{r warning = FALSE, fig.width=8, fig.height=5}


groups =6
#distance = "correlation" #euclidean
distance = "euclidean"

ph <- pheatmap(hm_dat[,selected_markers],cutree_rows = groups, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="ward")

# fn <- sprintf("%s_heatmap_6_tumor_%s.png", prefix, suffix)
#   ggsave(file.path(results_folder,  fn), ph, width = 6, height = 8)
```
#plot heatmap 
```{r warning = FALSE, fig.width=8, fig.height=8}


an$pgtu <- data.frame(cluster=rownames(hm_dat))  %>% 
  mutate(annotation = ifelse(cluster%in% c(1), "Inflamed_Glut1",
                        ifelse(cluster%in% c(2), "Inflamed_Ki-67",
                          ifelse(cluster%in% c(3,4), "Inflamed",
                            ifelse(cluster%in%c(8), "Excluded_SOX10",
                              ifelse(cluster%in% c(9), "Excluded", 
                                ifelse(cluster%in% c(13), "Mixed_SOX10",
                                  ifelse(cluster%in% c(7,12), "Mixed_Ki-67",
                                    ifelse(cluster%in% c(6), "Mixed_Glut1",
                                      ifelse(cluster%in% c(10), "Mixed_PDL1","Mixed"))))))))))

an_col$pgtu <- list(Cell_type=c("Inflamed_Glut1"=ts_pal$blue[1],
                              "Inflamed_Ki-67"=ts_pal$blue[2],
                              "Inflamed"=ts_pal$blue[3],
                              "Excluded_SOX10"=ts_pal$blue[4],
                              "Excluded"=ts_pal$blue[5],
                              "Mixed_SOX10"=ts_pal$blue[6],
                              "Mixed_Ki-67"=ts_pal$blue[7],
                              "Mixed_Glut1"=ts_pal$blue[8],
                              "Mixed_PDL1"=ts_pal$blue[9],
                              "Mixed"=ts_pal$blue[10]))

ph2 <- pheatmap(hm_dat[,selected_markers],cutree_rows = 1, clustering_distance_rows = distance,
  clustering_distance_cols = distance, method="complete", 
  border_color=NA, annotation_row = data.frame(Cell_type=an$pgtu$annotation), show_rownames = F, cellwidth = 20, cellheight = 15, 
  annotation_colors = an_col$pgtu, fontsize = 16, legend_breaks = seq(0,1,by=0.2))

fn2 <- sprintf("%s_heatmap_6_Tumor_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width =8 , height = 6)
```
#assigning  to PGtua
```{r}
m <- match(tmp$PGtu, an$pgtu$cluster)
tmp$PGtua <- an$pgtu$annotation[m]
sub$PGtua <- sub$PGa
sub[,tmp$id]$PGtua <- tmp$PGtua
table(sub$PGtua)
```

```{r}
ns <- table(tmp$sample_ROI_dsc_fig, tmp$PGtua)
soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(Sample=Var1, cell=Var2)
#re-order
soi_cf$cell<-factor(soi_cf$cell, levels=c("Inflamed_Glut1",
                              "Inflamed_Ki-67",
                              "Inflamed",
                              "Excluded_SOX10",
                              "Excluded",
                              "Mixed_SOX10",
                              "Mixed_Ki-67",
                              "Mixed_Glut1",
                              "Mixed_PDL1",
                              "Mixed"))

 
p <- ggplot(soi_cf, aes(Sample,Freq, fill = cell)) +
    scale_fill_manual("Cell type", values = ts_pal$blue, drop=FALSE) +geom_bar(stat = "identity") + 
  ggtitle('') + theme_light(base_size = 16)+
  theme(legend.title = element_text(face = "bold")) + labs(y ="Frequency")+
  scale_x_discrete(labels=c("Excluded", "Inflamed","Mixed\nexcluded","Mixed\ninflamed"))
  
p

  #save the plot
  fn <- sprintf("%s_frequency_6_Tumor_Sample_%s.pdf", prefix,suffix)
  ggsave(file.path(results_folder, fn), p, width = 6.5, height = 4)  

```




