---
title: "SABER melanoma data preprocessing"
output: html_notebook
---
## 0. Preparation
# Load packages

```{r}
library(S4Vectors)
library(SingleCellExperiment)
library(scater)
library(Rphenograph)
library(dplyr)
library(pheatmap)
library(RColorBrewer)
library(flowCore)
library(data.table)
```


# inputs
```{r}
panel_name <- "TH108_panel.csv"
roi_name <- "TH108_roiname.csv"
run=1
prefix = 'SABER_forMS'
suffix = paste0('r',run)

```

#parameters
```{r}
n_cells_dr = 500 # for t-sne

```


# Set the paths to folders and (meta)data
```{r}
folder.project <- dirname(getwd())
#folder for rds files
analysis_folder <- file.path(folder.project,"analysis")
ifelse(!dir.exists(analysis_folder), dir.create(analysis_folder), FALSE)
#folder for plots
results_folder <- file.path(folder.project,"analysis", "result_plots")
ifelse(!dir.exists(results_folder), dir.create(results_folder), FALSE)

#file path for (meta)data
fn.cpout <- file.path(folder.project,'output','cpout')
fn.cells <- file.path(fn.cpout,'cell.csv')
fn.image <- file.path(fn.cpout,'Image.csv')
fn.panel <- file.path(folder.project,'config', panel_name)
fn.meta <- file.path(fn.cpout,"acquisition_metadata.csv")
fn.roiname <-file.path(folder.project,'config',roi_name)
```

# Helpers

```{r helpers}
# construct data.frame for plotting with ggplot2
# using the specified assay slot as features values
.get_df <- function(sce, assay = "exprs") {
    dr <- do.call("cbind", reducedDims(sce))
    foo <- sapply(reducedDims(sce), ncol)
    colnames(dr) <- paste0(rep.int(names(foo), foo), sapply(foo, seq_len))
    df <- data.frame(dr, colData(sce), t(assay(sce, assay)), check.names = FALSE)
    reshape2::melt(df, id.vars = c(colnames(colData(sce)), colnames(dr)))
}
```

```{r, get matrix function }
.get_matrix <- function(sce,assay="scaled", group_by = "celltype", markers = rownames(sce)) {
  mat <- .get_df(sce[markers,cols=group_by], assay = assay) %>% mutate(group_by = get(group_by)) %>% 
    group_by(group_by, variable) %>%  summarize(median_scale_count = median(value))
  hm_dat = reshape2::dcast(data = mat, formula = 'group_by ~ variable',
                          value.var = 'median_scale_count')
  # save column
  trownames = hm_dat$group_by
  # convert to a matrix
  hm_dat = as.matrix(hm_dat[,-c(1)])
  row.names(hm_dat) = trownames
  return(hm_dat)
}
```


# Read in the cell data and metadata

```{r load-data}
cells <- read.csv(fn.cells , stringsAsFactors = FALSE)
image <- read.csv(fn.image, stringsAsFactors = FALSE)
meta <- read.csv(fn.meta, header = TRUE, stringsAsFactors = FALSE)
panel <- read.csv(fn.panel, header = TRUE, stringsAsFactors = FALSE)
roiname <- read.csv(fn.roiname, header = TRUE, stringsAsFactors = FALSE)
```

# Process data into a `SingleCellExperiment` object

```{r, processing-of-data}
# Select the features of interest
cur_cells <- cells[,grepl("MeanIntensity_FullStackFiltered_", colnames(cells))]
# adapt rownames to add the pannel
rownames(cur_cells) <- paste(cells$ImageNumber, cells$ObjectNumber, sep = "_")

# Create cell data
col.data <- DataFrame(row.names = rownames(cur_cells),
                       id = paste(cells$ImageNumber,cells$ObjectNumber,sep="_"),
                       ObjectNumber = cells$ObjectNumber,
                       ImageNumber = cells$ImageNumber,
                       X = round(cells$"Location_Center_X"),
                       Y = round(max(cells$"Location_Center_Y")-cells$"Location_Center_Y")
                       )
# Create panel data
row.data <- DataFrame(panel)
rownames(row.data) <- row.data$Target
# Order the features based on channels
channels.panel <- paste0("c", panel$channel)
channels.cells <- gsub(".*_", "", colnames(cur_cells))
cur_cells <- cur_cells[,match(channels.panel, channels.cells)]
# Scale the counts to incorporate the 16-bit specific scaling factor
cur_cells <- cur_cells*65536
```

#use the `SingleCellexperiment` package to handle and work with the data.
```{r, SCE-object}
# Generate the SCE object, 
sce <- SingleCellExperiment(assays = list(counts = t(cur_cells)))
colData(sce) <- col.data
rowData(sce) <- row.data
# Change rownames
rownames(sce) <- rownames(row.data)
```

#add the image metadata to the object.

```{r, image-meta}
sce$Image_path <- file.path(fn.cpout,image$FileName_CellImage[sce$ImageNumber])
sce$Image_folder <- fn.cpout
sce$Image_name <- image$FileName_CellImage[sce$ImageNumber]
sce$sample <- gsub(pattern = "(^.*?)_s0_p.*", replacement = "\\1", sce$Image_name)
sce$ROI <- gsub(pattern = "^.*_r(.*?)_.*", replacement = "\\1", sce$Image_name)
# Make matching "sample_ROIid" in both meta and colData
sce$sample_ROI <- paste(sce$sample,sce$ROI,sep = '_')
meta$sample_ROI <- paste(meta$AcSession,meta$AcquisitionID,sep = '_')
# Manually add ReDescription and AcSessionID in meta
sce$core_id <- meta$ReDescription[match(sce$sample_ROI, meta$sample_ROI)]
sce$sample_id <- meta$AcSessionID[match(sce$sample_ROI, meta$sample_ROI)]
sce$sample_ROI_id <- paste(sce$sample_id,sce$core_id,sep='_')
sce$prefix <- paste0(prefix,"_",suffix)
sce$cellnumber <- gsub(pattern = "^.*?_(.*)", replacement = "\\1", sce$id)
```
#checking colData
```{r}
unique(sce$sample_ROI)
unique(sce$ROI)
unique(sce$core_id)
unique(sce$sample_ROI_id)
unique(sce$sample_id)
unique(sce$sample)
unique(sce$prefix)
range(sce$X)
range(sce$Y)
```

# compute scaled expressions for visualization

#### to do: adapt cofactor for high intensity targets

```{r}
assay(sce, "exprs") <- asinh(assay(sce, "counts"))
```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(sce)), sce$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(sce[,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(sce, "TSNE") <- matrix(NA, ncol(sce), ncol(xy)) #create empty template
reducedDim(sce, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```

```{r}
table(sce$sample_id)
table(sce[,cs]$sample_ROI_id,sce[,cs]$sample_id)
```


#export fcs
```{r}

counts.matrix <- data.table(t(assay(sce[,cs],"counts")),type=sce[,cs]$sample_id)
coi <- rownames(sce)

#split the data by `Sample Name`
dt_list <- split(counts.matrix, by = c("type"))

names(dt_list)
#save fcs file 
fr_list <- lapply( 
  dt_list,
  function(x) {
    #data 
    dta <- as.matrix(x[,coi, with=FALSE])
    #meta
    meta <- data.frame(
      name = dimnames(dta)[[2]],
      desc = dimnames(dta)[[2]]
      )
    #meta$range <- apply(apply(dta,2,range),2,diff)
    #meta$minRange <- apply(dta,2,min)
    #meta$maxRange <- apply(dta,2,max)
    
    #meta$range <- c(0,4) #fixed range for cytobank gating, this doesnt work in cytobank
    ff <- new("flowFrame", exprs=dta, parameters=AnnotatedDataFrame(meta))
    write.FCS(ff, file.path(results_folder, paste0(paste0(x[, unique(type)]), ".FCS")))
  }
)

```


```{r prep-data}
saber_targets <- c("VISTA", "gp100", "CTLA4","PD-1","CD28","FoxP3","PDL2","SOX10","SOX9","CD4","CD8a", "TIM3", "LAG3", "CD86","CD11c", "PDL1")
assay(sce[saber_targets,], "exprs") <- asinh(assay(sce[saber_targets,], "counts")/5)
#assay(sce[c("PDL1","CD11c","SOX9" ),], "exprs") <- asinh(assay(sce[c("PDL1","CD11c","SOX9"),], "counts")/2)
es <- assay(sce, "exprs")
qs <- rowQuantiles(es, probs = c(0.0, 0.995))
x <- (es - qs[, 1]) / (qs[, 2] - qs[, 1])
x[x < 0] <- 0; x[x > 1] <- 1
assay(sce, "scaled") <- x
```

#show disribution of markers

###show TSNE cells

```{r, fig.width=20}
DR <- "TSNE"
var <- "scaled"
p <- .get_df(sce[,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =10) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'RdYlBu')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```
# color palette
```{r}
qual_col_pals <- brewer.pal.info[brewer.pal.info$category == 'qual',]   
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
pie(rep(1,length(col_vector)), col=col_vector, labels = names(col_vector))
```


```{r, fig.width=8}
DR <- "TSNE"
var <- "sample_ROI_id"
p <- .get_df(sce[,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = var)) +
    geom_point(alpha=1, size=1)+
    scale_color_manual(values=col_vector)+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```






##select ROIs for analysis
```{r}
selected_rois <- c('39765_t1','39765_t1-2','39765_t3','39765_t7','1936_t1','1936_t2','1936_t5','4324_t1','4324_t1-2','4324_t2','4324_t2-2','4324_t3','4324_t4','4324_t7')
sub <- sce [,sce$sample_ROI_id %in% selected_rois]
```

## Rename ROIs for figures
```{r}
sub$sample_ROI_id_fig <- roiname$sample_ROI_id_fig[match(sub$sample_ROI_id,roiname$sample_ROI_id)]
sub$sample_ROI_dsc_fig <- ifelse(sub$sample_id %in% c("1936"), "Excluded" , 
                                   ifelse(sub$sample_id %in% c("39765"), "Inflamed" ,
                                   ifelse(sub$sample_ROI_id %in% c("4324_t3","4324_t7"), "Mixed_Excluded" ,"Mixed_Inflamed")))
unique(sub$sample_ROI_id_fig)
unique(sub$sample_ROI_dsc_fig)
```


```{r}
table(sub$sample_ROI_id)
```


## 1. Clustering 1: detect unspecific bright spots
# Phenograph clustering 
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
selected_markers <- c('FoxP3','SOX10','LAG3','CD86')
pg_res <- Rphenograph(data = t(assay(sub[selected_markers,],"exprs")), k = 20)
sub$PGu <- factor(pg_res[[2]]$membership)
```

```{r, fig.width=6, fig.height=12}
hm_dat <- .get_matrix(sub, group_by="PGu")
groups =2

clust_distance = "euclidean"
clust_method = "ward"

pheatmap(hm_dat[,selected_markers], clustering_distance_cols = clust_distance,
         clustering_distance_rows = clust_distance,
         clustering_method = clust_method)
```

```{r}
an <- data.frame(Annotation = ifelse(rownames(hm_dat)==8, "Removed", "Kept"))
ann_colors = list(Annotation = c(Kept="navy", Removed="firebrick2"))


ph2 <- pheatmap(hm_dat[,selected_markers],
                clustering_distance_cols = clust_distance,
         clustering_distance_rows = clust_distance,
         clustering_method = clust_method,
  border_color=NA, annotation_row = an, show_rownames = F, cellwidth = 20, cellheight = 10, fontsize = 16,
  annotation_colors = ann_colors)

fn2 <- sprintf("%s_heatmap_1_usp_%s.pdf", prefix, suffix)
  ggsave(file.path(results_folder,  fn2), ph2, width = 6, height = 8)
```

##TSNE plot to show outlier


#assigning to sub$PGua
```{r}
usp <- c(8)
sub$PGua <- ifelse(sub$PGu %in% usp, "usp","non_usp")
```

#plot cell frequencies
```{r}

ann_colors$usp <- c(non_usp="navy", usp="firebrick2")

ns <- table(sub$sample_ROI_id_fig, sub$PGua)

soi_cf <- data.frame(prop.table(ns, 1))%>%
  rename(ROI=Var1, cell=Var2)

  p <- ggplot(soi_cf, aes(ROI, Freq, fill = cell)) +
    scale_fill_manual("cell type", values = ann_colors$usp, drop=FALSE) +
    geom_bar(stat = "identity") + ggtitle('') + theme_light()
  p

  #save the plot
  fn <- sprintf("%s_frequency_1_usp_%s.pdf", prefix,suffix)
  ggsave(file.path(results_folder, fn), p, width = 12, height = 4)  
```
#remove usp
```{r}
sub <- sub[,sub$PGua %in% c("non_usp")]
```

# compute scaled expressions for visualization
```{r}
# assay(sub, "exprs") <- asinh(assay(sub, "counts"))
es <- assay(sub, "exprs")
qs <- rowQuantiles(es, probs = c(0.0, 0.999))
x <- (es - qs[, 1]) / (qs[, 2] - qs[, 1])
x[x < 0] <- 0; x[x > 1] <- 1
assay(sub, "scaled") <- x
```

```{r}
rownames(sub)
```

# Run tSNE
```{r}
set.seed(200, kind = "Mersenne-Twister", normal.kind = "Inversion"); rnorm(1)
# split cells by sample
cs <- split(seq_len(ncol(sub)), sub$sample)
# sample 'n_cells_dr' per sample
cs <- unlist(lapply(cs, function(u) 
  sample(u, min(n_cells_dr, length(u)))))

# run 'TSNE' dimension reductions on selected channels 
xy <- calculateTSNE(sub[,cs], exprs_values = "exprs",perplexity = 50)
reducedDim(sub, "TSNE") <- matrix(NA, ncol(sub), ncol(xy)) #create empty template
reducedDim(sub, "TSNE")[cs, ] <- xy #add TSNE coordinates
rm(xy)
```



#show disribution of markers

###show TSNE cells

```{r, fig.width=20}
DR <- "TSNE"
var <- "scaled"
p <- .get_df(sub[,cs],assay="scaled") %>%
    ggplot(aes_string(paste0(DR,"1"), paste0(DR,"2"), col = "value")) +
    geom_point(alpha=1, size=1)+ 
    facet_wrap("variable", ncol =10) +
    scale_color_gradientn(colours=rev(brewer.pal(11, 'Spectral')))+
    theme_void() + theme(aspect.ratio = 1, strip.text.x = element_text(size = 8))

print(p)

#fn <- sprintf("%s__%s_%s.png", pp_prefix, var,DR)
  #ggsave(file.path(vsrs_folder,  fn), p, width = 24, height = 18)
  
```

